---
title: "Women Educational Attainment and Childbearing"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: flatly
    source_code: embed
    embed-resources: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(rio)
library(ggplot2)
library(tidyverse)
library(here)
library(gtsummary)
library(kableExtra)
library(plotly)
library(gapminder)
library(ggtext)
library(gganimate)
library(magick)
library(dplyr)
library(tidyr)
library(viridis)
```

```{r ,include=FALSE}
woman <- import(here("data","BL_v3_F2564.csv"))%>%
  as.tibble()

child <- import(here("data","childbearing.csv"))%>%
  as.tibble()

child$country <- child$`Region, subregion, country or area *`
child$year <- child$Year

### combine dataset

d <- left_join(woman, child, join_by(country, year))

d <- d[, -c(21, 22)]

d$childbearing <- as.numeric(d$`Mean Age Childbearing (years)`)

d<- d %>%
  group_by(country) %>%
  mutate(Mean.year.of.childbearing = mean(childbearing))

head(d)

### rename variables

d <- d %>%
   rename("No_formal" = "lu", 
          "Primary" = "lp",
          "Primary_complete" = "lpc",
          "Secondary" = "ls",
          "Secondary_complete" = "lsc",
          "% of Tertiary" = "lh",
          "Tertiary_complete" = "lhc",
          "Schooling.year" = "yr_sch",
          "Average.tertiary.schooling" = "yr_sch_ter",
          "Region" = "region_code",
          "Population" = "pop"
          )

head(d)
sapply(d, function(x) sum(is.na(x))) 
dclean <- d %>%
  drop_na()

```

# About this project

## Column {data-width="400"}

### Background

Exploring the patterns of women's educational attainment and childbearing and their relationship is crucial for understanding broader societal dynamics and promoting gender equality. Education empowers women with knowledge, skills, and the confidence to make informed decisions about their lives, including family planning and career choices. Higher levels of education often correlate with later ages of marriage and childbearing, smaller family sizes, and improved maternal and child health outcomes. This, in turn, contributes to better economic prospects for women and their families, reduces poverty, and promotes sustainable development. Moreover, understanding these patterns helps policymakers and stakeholders design more effective educational and health programs that address the unique needs of women across different cultures and societies. By investing in women's education, societies can break cycles of poverty, improve quality of life, and foster environments where gender equality can thrive.

**Description of Data and Source**

The data I used for this final project come from two datasets. The first one is the Barro-Lee dataset reported in Barro and Lee (2013), including 2,044 observations on 20 variables. The data reflect Barro-Lee estimates of educational attainment for the female population aged 25-64 years from 1950 to 2015. This dataset is publicly available from the website: <http://www.barrolee.com/>. More information related to the detailed data sources and missing observations estimation were described in the original paper. Another dataset is the United Nations World Population Prospects (2022) dataset available from <https://population.un.org/wpp/Download/>, which has 20,596 observations on 65 variables including the one of interest _Mean Age Childbearing (years)_ from 1950 to 2100. In order to answer the research questions of this project, I extracted data of _Mean Age Childbearing (years)_ in 1950-2015 and joined it to the Barro-Lee dataset.

**Research Questions**

This project focuses on four questions:

1.  What is the relationship between the mean age of first childbearing and the average total years of schooling among the female population across different countries between 1950 and 2015?

2.  How has the relationship between the population and the average years of tertiary education for women in South Asian countries changed from 1950 to 2015?

3.  What is the trend in the total years of schooling for women in South Asian countries from 1950 to 2015?

4.  How has the total years of schooling at different educational levels for women in the US evolved from 1950 to 2015?

## Column {data-width="400"}

### Descriptive Table

```{r}
d1 <- dclean[, -2]

table <- d1 %>% 
  select(6:16, 19, 22) %>%
  tbl_summary(
  statistic = list(
    all_continuous() ~ "{mean}({sd})",
    all_categorical() ~ "{p}%"
    ), 
    label = list(
    No_formal ~ "No formal education",
    Primary ~ "Primary as highest level",
    Primary_complete ~ "Primary completed",
    Secondary ~ "Secondary as highest level",
    Secondary_complete ~ "Secondary completed",
    "% of Tertiary" ~ "Tertiary as highest level",
    Tertiary_complete ~ "Tertiary completed",
    Schooling.year ~ "Average years of total schooling",
    yr_sch_pri ~ "Average years of primary schooling",
    yr_sch_sec ~ "Average years of secondary schooling",
   Average.tertiary.schooling= "Average years of tertiary schooling",
    Mean.year.of.childbearing ~ "Mean age of first childbearing",
    Population ~ "female population (in 1000s)"
   )
) %>% 
  modify_caption("**Table 1. Descriptive Statistics**")

table

```

In this combined dataset, there are 1,862 observations for 23 variables. Based on the information from Table 1, the average global female population in 1950-2015 was 6.775million. Between the age of 28-29 years (Mean = 28.72, SD = 1.15), the female would have their first child. As for educational attainment, 38% of the global female population didn’t receive formal education in 1950-2015. Meanwhile, across all countries within the sample, 31% of the female population had primary school as their highest level of schooling, while 16% had it completed. Twenty four percent of the sample had secondary school as their highest level, while 14% completed it. Only 7% entered tertiary level of schooling, and only 4.6% had it finished. The average years of total schooling of female population were 5.1 years (SD = 3.7) across the world in 1950-2015, while the average years of primary level were 3.3 years (SD = 2.16), along with 1.61 years (SD = 1.64) for the secondary level and 0.24 years (SD = 0.34) for the tertiary level. Besides, in this sample, 17% of countries belong to the group of Advanced Economies, 12% as East Asia and the Pacific, 14% as Europe and Central Asia, 17% as Latin America and the Caribbean, 13% as Middle East and North Africa, 5.3% as South Asia, and 23% as Sub-Saharan Africa.

# Vis 1

## Inputs {.sidebar}

**A story of this visualization**

This visualization aims to elucidate the relationship between educational attainment and the mean age of childbearing for women globally from 1950 to 2015, addressing the first research question. It incorporates variables such as the average years of total schooling, the mean age of childbearing, country names, region codes, the female population, and the year. Here, 'the average years of total schooling' denotes the average length of education received at primary, secondary, or tertiary institutions. 'The mean age of childbearing' is the average age at which women have their first child. 'Population' specifically refers to the female population of each country.

The analysis reveals an unexpected negative correlation between educational attainment and the mean age of childbearing for women worldwide during the period 1950-2015. A linear regression model indicates that an increase of one unit in the average years of total schooling correlates with a decrease of 1.26 units in the mean age of childbearing, on average, among the female population. However, it's important to recognize that this relationship may vary across different regions or countries, necessitating further investigation. 

**The history and evolution of the visualization**

In the original version, I utilized data solely from 2015, with data points colored by region (e.g., advanced economies, South Asia) and differentiated point sizes based on the female population of each country. Additionally, I set the y-axis to start at 0. In the second version, I enhanced the plot by adding a linear model fit line to demonstrate the relationship between educational attainment and average childbearing age. To ensure the data remained visible, I chose not to include the confidence interval.

After integrating feedback for the final version, I adjusted the y-axis to begin at 24, aiming to highlight a more distinct pattern within the data. Furthermore, I also incorporated a linear regression fit line to depict the correlation between women's educational attainment and the mean age of childbearing spanning from 1950 to 2015.

**Audience of this visualization**

This visualization is designed for an audience with limited statistical background knowledge.


## Row {data-width="260"}

### Final version

```{r, fig.width = 8, fig.height = 5 }

model1 <- lm(Schooling.year ~ Mean.year.of.childbearing, data = dclean)
intercept1 <- coef(model1)[1]
slope1 <- coef(model1)[2]
equation1 <- paste("y = ", round(slope1, 2), "x", " + ", round(intercept1, 2))


p1 <- dclean %>%
  ggplot(aes(x=Schooling.year, y=Mean.year.of.childbearing, label = country)) + 
 geom_point(aes(color = Region, frame = year, size = Population)) +
  scale_color_viridis(discrete = TRUE)+
    annotate("text", x = 12, y = 24.5, label = equation1, hjust = 0, vjust = 0, color = "black") +
  geom_smooth(method = lm, se = FALSE) + 
  scale_x_continuous("Average years of total schooling", 
                     breaks = seq(0, 14,2),
                            limits = c(0,14)) +
  scale_y_continuous("Mean age of first childbearing",
                    breaks = seq(24, 32,4),
                         limits = c(24,32)) +
   theme_minimal(base_size = 10) +
  labs(
    title = "A negative relationship was found between educational attainment and \n the mean age of childbearing for women worldwide (1950-2015)",
    subtitle = "An increase of one unit in the educational attainment correlates with \n a decrease of 1.26 units in the mean age of childbearing",
    x = "Total years of schooling",
    y = "Mean age of childbearing",
    color = "Region",
    size = "",
    caption = "Note: The sizes of the data points represent population sizes. \n Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/"
  ) +
theme(
  plot.title = element_text(size = 10),
  legend.position = "right",
  legend.direction = "vertical", 
  legend.title = element_text(size = 8), 
  legend.text = element_text(size = 6), 
  axis.ticks.x = element_line(color = "black"),
  axis.text.x = element_text(vjust = 5),
  axis.ticks.length = unit(-0.5, "cm"),
  axis.line = element_line(color = "black"),
  panel.grid.major = element_blank()
)


p1_plotly <- ggplotly(p1)

p1_plotly <- p1_plotly %>%
  layout(annotations = list(
    text = 'Note: The sizes of the data points represent population sizes. \n Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/',
     x = 1.03,
    y = -0.125, 
    showarrow = FALSE,
    xref = 'paper',
    yref = 'paper',
    xanchor = 'right',
    yanchor = 'bottom',
    font = list(
      size = 5
  )))

p1_plotly

```

## Column {.tabset data-width="220,"}

### Original version

```{r, fig.width = 4, fig.height = 6}

p1 <- dclean %>%
  filter(year == 2015) %>%
  ggplot(aes(x=Schooling.year, y=Mean.year.of.childbearing, color = Region)) + 
 geom_point(aes(size = Population)) +
  scale_x_continuous("Average years of total schooling", 
                     breaks = seq(0, 14,2),
                            limits = c(0,14)) +
  scale_y_continuous("Mean age of first childbearing",
                    breaks = seq(0, 34,4),
                         limits = c(0,34)) +
   theme_minimal(base_size = 10) +
  labs(
    title = "Relation between the educational attainment and mean age of <br> childbearing for female population worldwide in 2015",
    x = "Total years of schooling",
    y = "Mean age of childbearing",
    color = "Region",
    size = "Population shown in point size",
    caption = "Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/"
  ) +
theme(
  plot.title = element_markdown(margin = margin(b = 5), size  = 12),
  legend.position = "right",
  legend.direction = "vertical", 
  legend.title = element_text(size = 8), 
  legend.text = element_text(size = 6), 
  axis.line.x = element_line(color = "black"), 
  axis.line.y = element_line(color = "black"),
  panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_blank(),
  plot.caption       = element_text(hjust = 1, 
                                              size  = 6,
                                              margin=margin(t=5))
)

ggplotly(p1)

```

### Second version

```{r, fig.width = 10, fig.height = 6 }

d2 <- dclean %>%
  filter(year == 2015) 

model <- lm(Schooling.year ~ Mean.year.of.childbearing, data = d2)

intercept <- coef(model)[1]
slope <- coef(model)[2]

equation <- paste("y = ", round(slope, 2), "x", " + ", round(intercept, 2))


d2 %>%
  ggplot(aes(x=Schooling.year, y=Mean.year.of.childbearing, label = country)) + 
 geom_point(aes(size = Population, frame = year, color = Region)) +
  geom_smooth(method = lm, se = FALSE) + 
   annotate("text", x = 0, y = 32, label = equation, hjust = 0, vjust = 0, color = "red") +
  scale_x_continuous("Average years of total schooling", 
                     breaks=seq(0, 14, by = 2),
                     limits = c(0, 14)) +
  scale_y_continuous("Mean age of first childbearing",
                     breaks = seq(0, 34, by = 4),
                     limits = c(0, 34)) +
   theme_minimal(base_size = 10) +
  labs(
    title = "Relation between the educational attainment and mean age of \n childbearing  for female population worldwide in 2015",
    x = "Total years of schooling",
    y = "Mean age of childbearing",
    color = "Region",
    size = "Population shown in point size",
    caption = "Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/"
  ) +
theme(
  plot.title = element_markdown(size  = 10),
  legend.position = "right",
  legend.direction = "vertical", 
  legend.title = element_text(size = 8), 
  legend.text = element_text(size = 6), 
  axis.line.x = element_line(color = "black"), 
  axis.line.y = element_line(color = "black"),
  panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_blank()
)+
  scale_y_continuous(breaks = seq(0, 34,4),
                         limits = c(0,34))+
  scale_x_continuous(breaks = seq(0, 14,2),
                            limits = c(0,14))

```

# Vis 2

## Inputs {.sidebar}

**A story of this visualization**

The purpose of this visualization is to illustrate the relationship between the female population and the average years of tertiary education across countries in South Asia from 1950 to 2015. The variables presented include population, average years of tertiary schooling, country names, the percentage of the population with tertiary education as their highest level of educational attainment, and the year. The analysis of the plot reveals that, for Bangladesh, India and Pakistan, there is a general upward trend in both the average years of tertiary education and the percentage of population with tertiary schooling as the highest education level. In terms of Sri Lanka, the peak in both variables was observed in 2010, followed by a significant decline in 2015. For Nepal, Afghanistan, and Maldives, an obvious increase was observed only in the average years of tertiary schooling. 

**The history and evolution of the visualization**

In the original version, I color-coded data points by country and varied point sizes according to the percentage of the female population with tertiary education as their highest level of educational attainment.

In the final version, I enhanced the plot to be interactive, showcasing data from 1950 to 2015 and including pop-up windows that appear when viewers hover over data points. Recognizing that population growth rates can be exponential, I also incorporated a panel that converts the x-axis to a logarithmic scale for a more nuanced interpretation of the data.

**Audience of this visualization**

The original panel of the visualization is designed for the general public, while the logarithmic panel is intended for an audience with a statistical background.


## Column {data-width="280"}

### Final version

```{r, fig.width = 10, fig.height = 6 }

d3 <- dclean %>%
  filter(Region == "South Asia" )

plot2 <-  d3 %>%
  ggplot(aes(x=Population, y = Average.tertiary.schooling, color = country)) + 
  geom_point(aes(size = `% of Tertiary`, frame = year)) +
    scale_color_viridis(discrete = TRUE)+
   theme_minimal(base_size = 10) +
  labs(
    title = "The population of women and the average years of tertiary schooling \n both increased significantly in South Asian countries (1950-2015)",
    x = "Population (in 1000s)",
    y = "Average years of tertiary schooling",
    color = "Country Name",
    size = "",
    caption = "Note: The sizes of the data points represent population sizes. \n Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/"
  ) +
theme( plot.title = element_text(size = 10),
legend.position = "right", 
legend.direction = "vertical", 
legend.title = element_text(size = 8),
legend.text = element_text(size = 8), 
panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_blank(),
axis.line.x = element_line(color = "black"), 
  axis.line.y = element_line(color = "black"),
) +
  scale_x_continuous(breaks = seq(0, 400000,50000),
                         limits = c(0,400000),
                     labels = scales::comma)+
  scale_y_continuous(breaks = seq(0, 0.7,0.1),
                            limits = c(0,0.7))

p2 <- ggplotly(plot2) 

# Define button for switching to logarithmic scale
to_log_button <- list(
  method = "relayout",
  args = list(list(
    xaxis = list(
      type = "log", 
      range = list(log10(10), log10(400000)),
      title = "Log of population",
      titlefont = list(size = 10), 
      tickfont = list(size = 10)
    ), 
    yaxis = list(
      type = "linear", 
      range = c(0,0.7),
      titlefont = list(size = 10), 
      tickfont = list(size = 10),
      title = "Average years of tertiary schooling",
      showline = TRUE,  
      linecolor = "black", 
      linewidth = 0.5,  
      dtick = 0.1
    )
  )),
  label = "Log"
)

# Button to reset to the original plot (linear scale in this case)
reset_to_original <- list(
  method = "relayout",
  args = list(list(
    xaxis = list(
      type = "linear",
      range = c(0, 400000),
      autorange = FALSE,
      title = "Population (in 1000s)",
      titlefont = list(size = 10),
      tickfont = list(size = 10), 
      tickformat = ",", 
      tickmode = "auto",
      nticks = 20
    ),
    yaxis = list(
      type = "linear",
      range = c(0, 0.7),
      autorange = FALSE,
      title = "Average years of tertiary schooling",
      titlefont = list(size = 10), 
      tickfont = list(size = 10), 
      dtick = 0.1 
    )
  )),
  label = "Reset"
)

p2 <- layout(p2, updatemenus = list(
  list(
    type = "buttons",
    direction = "vertical",
    xanchor = 'right',
    yanchor = "top",
    pad = list(r= 0, t= 5, b = 5),
    x = 1,
    y = 1,
    buttons = list(reset_to_original, to_log_button) # Order can be adjusted
  )
))

p2.1 <- layout(p2, 
             annotations = list(
               list(
                 text = "Note: The sizes of the data points represent population sizes. <br>Source: Barro and Lee (2013), http://www.barrolee.com/ <br>UN, World Population Prospects, https://population.un.org/wpp/Download/",
                 x = 1.03,
                 y = -0.085, 
                 showarrow = FALSE,
                 xref = 'paper',
                 yref = 'paper',
                 xanchor = 'right',
                 yanchor = 'bottom',
                 font = list(
                   size = 6
                 )
               )
             ),
             margin = list(b = 100) # Adjust bottom margin to ensure caption is visible
            )

p2.1

```

## Column {.tabset data-width="220,"}

### Original version

```{r, fig.width = 10, fig.height = 6 }

d3.1 <- dclean %>%
  filter(year ==2015) %>%
  filter(Region == "South Asia" )

plot2 <-  d3.1 %>%
  ggplot(aes(x=Population, y = Average.tertiary.schooling, color = country)) + 
  geom_point(aes(size = `% of Tertiary`)) +
   theme_minimal(base_size = 10) +
  labs(
    title = "Relation between the female population and the average years of \n tertiary schooling for countries in South Asia in 2015",
    x = "Population (in 1000s)",
    y = "Average years of tertiary schooling",
    color = "Country Name",
    size = "% of tertiary shown in point size",
    caption = "Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/"
  ) +
theme(plot.title = element_markdown(size  = 10),
legend.position = "right", # Position legend at the bottom
legend.direction = "vertical", # Orient legend horizontally
legend.title = element_text(size = 8), 
legend.text = element_text(size = 8), 
axis.line.x = element_line(color = "black"), 
  axis.line.y = element_line(color = "black"),
) +
  scale_x_continuous(breaks = seq(0, 400000,50000),
                         limits = c(0,400000),
                     labels = scales::comma)+
  scale_y_continuous(breaks = seq(0, 0.35,0.05),
                            limits = c(0,0.35))

plot2

```


# Vis 3

## Inputs {.sidebar}

**A story of this visualization**

The purpose of this visualization is to illustrate the trend in the total years of schooling for women in South Asian countries from 1950 to 2015. The variables included in this visualization are the average years of total schooling, country names, and the year. The plot reveals that in countries like Afghanistan, Bangladesh, India, Nepal, and Pakistan, there has been a consistent rise in the average years of total schooling from 1950 to 2015, starting from less than two years. For the other two countries, the Maldives and Sri Lanka, the average years of educational attainment began at a level higher than two years. Sri Lanka has experienced a significant, accelerated improvement in educational attainment for women. Before 1985, the Maldives showed a steady increase in average schooling for the female population, which was followed by a sharp decline until 2000 and then a reversal to an increasing trend up until 2015. 

**The history and evolution of the visualization**

The original version featured data lines color-coded by country, with data points for every year and a separate legend for illustration. In the second version, I transformed the plot into a dynamic format. However, noticing that the patterns of change in Maldives and Sri Lanka differed from other countries, I developed the final version to highlight these distinctions.
Based on feedback, in this final version, I eliminated the x and y axes lines, ensured there was space between the x-axis line and the x-axis labels, and annotated each line with country names, thereby removing the need for a legend. Additionally, I corrected the y-axis title to ‘Average Years of Total Schooling’. 

**Audience of this visualization**

This visualization is designed for the general public.



## Column {data-width="400"}

### Final version

```{r, fig.width = 8, fig.height = 6 }

d4 <- dclean %>%
  filter(Region == "South Asia") %>%
  select(country, year, Schooling.year, yr_sch_sec)

p3 <- ggplot(d4, aes(x = year, y = Schooling.year, group = country, color = country)) +
    geom_point(data=d4,aes(x=year,y=Schooling.year), shape =21, color = "black", fill = "white", size=3) +
  geom_path()+
  theme_minimal() +
  labs(
       x = "Year",
       y = "Average years of total schooling",
       color = "Country",
       caption = "Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/") +
  theme(
legend.position = "right", 
legend.direction = "vertical", 
legend.title = element_text(size = 10), 
legend.text = element_text(size = 6), 
  axis.line = element_blank(),
panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.text.x = element_text()
) +
  scale_x_continuous(breaks = seq(1950, 2015,5),
                         limits = c(1950,2015))+
  scale_y_continuous(breaks = seq(0, 12,2),
                            limits = c(0,12))+
  transition_reveal(year) 

my_color          <- rep('gray',length(unique(d4$country)))
  names(my_color)   <- unique(d4$country)
  my_color['Maldives'] <- 'blue'
 my_color['Sri Lanka'] <- 'orange'


labels <- subset(d4,country = c('India', 'Maldives'))
 
p4<-  p3 +
  labs(title = "Significant Increase in Average Years of Total Schooling for Women in <span style='color:orange'>Sri Lanka </span> Versus Temporary Decrease in <span style='color:blue'>Maldives </span> (1950-2015)") +
  theme(plot.title = element_markdown(size  = 10))+
   scale_color_manual(values = my_color)+
    geom_text(data=labels,aes(x=year,y=Schooling.year,label=country),
              size = 4,hjust=-0.2,vjust=1)+
    guides(size=FALSE,color=FALSE)

animate(p4, 
          duration = 23, 
          fps = 20, 
          width = 800, 
          height = 600, 
          renderer = gifski_renderer(),
         end_pause = 10)



```

## Column {.tabset data-width="225,"}

### Original version

```{r}



d4.1 <- dclean %>%
  filter(Region == "South Asia") %>%
  select(country, year, Schooling.year, yr_sch_sec)

p3 <- ggplot(d4.1, aes(x = year, y = Schooling.year, group = country, color = country)) +
    geom_point(data=d4.1,aes(x=year,y=Schooling.year), shape =21, color = "black", fill = "white", size=3) +
  geom_path(aes(group = country, frame = year)) +
  theme_minimal() +
  labs(title = "Changes in total schooling by country in South Asia (1950-2015)",
       x = "Year",
       y = "Total years of schooling",
       color = "Country",
       caption = "Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/"
       ) +
  theme(
legend.position = "right", 
legend.direction = "vertical", 
legend.title = element_text(size = 10), 
legend.text = element_text(size = 6), 
axis.line.y = element_line(color = "black"), 
  axis.line.x = element_blank(),
panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.text.x = element_text(vjust = 10)
) +
  scale_x_continuous(breaks = seq(1950, 2015,5),
                         limits = c(1950,2015))+
  scale_y_continuous(breaks = seq(0, 12,2),
                            limits = c(0,12))+
     geom_hline(yintercept = 0, color = "black", size = 0.2)

p3

```

### Second version

```{r}


p3 <- ggplot(d4.1, aes(x = year, y = Schooling.year, group = country, color = country)) +
    geom_point(data=d4.1,aes(x=year,y=Schooling.year), shape =21, color = "black", fill = "white", size=3) +
  geom_path()+
  theme_minimal() +
  labs(title = "Changes in total schooling by country in South Asia (1950-2015)",
       x = "Year",
       y = "Total years of schooling",
       color = "Country",
       caption = "Source: Barro and Lee (2013), http://www.barrolee.com/ \n UN, World Population Prospects, https://population.un.org/wpp/Download/") +
  theme(
legend.position = "right",
legend.direction = "vertical", 
legend.title = element_text(size = 10), 
legend.text = element_text(size = 6),
axis.line.y = element_line(color = "black"), 
  axis.line.x = element_blank(),
panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.text.x = element_text(vjust = 10)
) +
  scale_x_continuous(breaks = seq(1950, 2015,5),
                         limits = c(1950,2015))+
  scale_y_continuous(breaks = seq(0, 12,2),
                            limits = c(0,12))+
     geom_hline(yintercept = 0, color = "black", size = 0.2) +
  transition_reveal(year) 

 animate(p3, 
          duration = 20, 
          fps = 20, 
          width = 800, 
          height = 600, 
          renderer = gifski_renderer())


```


# Vis 4

## Inputs {.sidebar}

**A story of this visualization**

For this visualization, I crafted a GIF to dynamically illustrate how the total years of schooling across different educational levels for women in the U.S. evolved from 1950 to 2015. It is apparent that the average duration of primary education has seen little change post-1975, whereas the average duration of tertiary education has experienced a considerable rise during this 65-year timeframe. Unfortunately, even though I tried different approaches, I didn't succeed in calibrating the y-axes to uniform ranges, which might impact the readability of the plot. 

**Audience of this visualization**

This visualization is designed for the general public.


## Column {data-width="400"}

### Final version

```{r}

dir_out <- file.path(tempdir(), "US")
dir.create(dir_out, recursive = TRUE)

US<- d %>%
  filter(country == "USA") %>%
  select(year, Schooling.year, yr_sch_pri, yr_sch_sec, Average.tertiary.schooling)

US <- US %>%
  rename("Total years of schooling" = "Schooling.year",
         "Years of primary schooling" = "yr_sch_pri",
         "Years of secondary schooling" = "yr_sch_sec",
         "Years of tertiary schooling" = "Average.tertiary.schooling")

US <- US %>%
  pivot_longer(cols = c(3:6),
    names_to = "School_level",
    values_to = "Years_of_schooling")

level <-
  US %>%
  pull("School_level") %>%
  unique(.) %>%
  sort(.)

for (n in level) {
plot4 <-
    US %>%
  filter(School_level == n) %>%
    ggplot(aes(year,Years_of_schooling)) +
    geom_col()+
    theme_minimal()+
  labs(caption = "Note: Changes in Educational Attainment at Different Levels for Women in the US (1950-2015)")+
  theme(panel.grid.major.y = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.minor.x = element_blank(),
  axis.line.y = element_line(color = "black"),
  axis.ticks.y = element_line(color = "black"),
  axis.text.x = element_text(vjust = 8))+
  scale_x_continuous(breaks = seq(min(US$year), max(US$year), by = 5))+
   scale_y_continuous(breaks = seq(0, max(US$Years_of_schooling), by = 2))+
   geom_hline(yintercept = 0, color = "black", size = 0.5)+
  labs(x = "Year", y = "Average years of schooling", title = n)

fp <- file.path(dir_out, paste0(n, ".png"))
  
  ggsave(plot = plot4, 
         filename = fp, 
         device = "png")

}


## list file names and read in
imgs <- list.files(dir_out, full.names = TRUE)
img_list <- lapply(imgs, image_read)

## join the images together
img_joined <- image_join(img_list)

## animate at 0.5 frames per second
img_animated <- image_animate(img_joined, fps = 0.5)

## view animated image
img_animated

## save to disk
image_write(image = img_animated,
            path = "school.gif")



  
```
